A tutorial on spatiotemporal partially observed Markov process models via the R package spatPomp

### 2024-08-07 ms.Rnw
add R script and recompute with R 4.4.1

<<bm_rho_prof_eval,eval=T,cache=F,echo=F>>=
rho_prof <- bake(file = paste0(bm_files_dir,"rho-profile.rds"),{
<<bm_rho_prof_code>> 
})
@
Above, calling \code{igirf} on \code{ig1} imports all the previous algorithmic settings except for those that we explicitly modify.
Each row of \code{rho\_prof} now contains a parameter estimate its log-likelihood, with $\rho$ values fixed along a grid.
The MCAP 95\% confidence interval constructed by \code{mcap} uses \code{loess} to obtain a smoothed estimate of the profile likelihood function and then determines a confidence interval using by a cutoff based on the delta method applied to a local quadratic regression.
This cutoff is typically slightly larger than the asymptotic 1.92 cutoff for a standard profile likelihood confidence interval constructed assuming error-free likelihood maximization and evaluation.
<<bm_rho_prof_mcap>>=
rho_mcap <- mcap(rho_prof[,"loglik"],parameter=rho_prof[,"rho"])
rho_mcap$ci
@
Note that the data in \code{bm10} are generated from a model with $\rho=0.4$.

>>>>

<<bm_rho_prof_eval,eval=T,cache=F,echo=F>>=
rho_prof <- bake(file = paste0(bm_files_dir,"rho-profile.rds"),{
<<bm_rho_prof_code>>
attr(rho_prof,"cores") <- cores
rho_prof
})
@
Above, calling \code{igirf} on \code{ig1} imports all the previous algorithmic settings except for those that we explicitly modify.
Each row of \code{rho\_prof} now contains a parameter estimate its log-likelihood, with $\rho$ values fixed along a grid.
The MCAP 95\% confidence interval constructed by \code{mcap} uses \code{loess} to obtain a smoothed estimate of the profile likelihood function and then determines a confidence interval using by a cutoff based on the delta method applied to a local quadratic regression.
This cutoff is typically slightly larger than the asymptotic 1.92 cutoff for a standard profile likelihood confidence interval constructed assuming error-free likelihood maximization and evaluation.
<<bm_rho_prof_mcap>>=
rho_mcap <- mcap(rho_prof[,"loglik"],parameter=rho_prof[,"rho"])
rho_mcap$ci
@
Note that the data in \code{bm10} are generated from a model with $\rho=0.4$.
Even for this toy example, the profile likelihood confidence interval requires a non-trivial computational effort.
This calculation took $\Sexpr{myround(attr(rho_prof,"system.time")["elapsed"]/60/60,1)}$ hours using $\Sexpr{attr(rho_prof,"cores")}$ cores.
As a side-effect, the many searches involved in obtaining a profile likelihood ensure that the likelihood surface is extensively explored.
